Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-02-25 manuscript.Rmd
close to the end of the intro

Classification accuracy from ML is good. older papers; Airaksinen; Frontiers

What's missing

>>>>

Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults [CITEs], children [CITEs], and infants [CITEs]. Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [CITE] and OTHER EXAMPLES? Our openly-shared dataset [CITE] contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.

REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.

### 2023-02-25 manuscript.tex
close to the end of the intro

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session {[}CITE ADOLPH, ANYONE ELSE?{]}, but sometimes a fixed camera may be used {[}IVERSON?{]}. The primary advantage of video recording is that it captures real-time behavior. Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior. However, requiring an experimenter to pan a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior). Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine. Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants. Our annotation of 90 minutes of body position takes approximately XX-XX minutes to complete (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) would take XX hours to annotate.

@@

Wear time is not an issue. Lots of Smith and other studies with long recordings.

Classification accuracy from ML is good. older papers; Airaksinen; Frontiers

What's missing

>>>>

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session {[}CITE ADOLPH, ANYONE ELSE?{]}, but sometimes a fixed camera may be used {[}IVERSON?{]}. The primary advantage of video recording is that it captures real-time behavior. Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior. However, requiring an experimenter to pan a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior). Typical video recording sessions last XX-XX, short of capturing the variety of activities across the full daily routine. Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants. For annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to full annotate.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home. Lightweight sensors (XX-XXg) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior. Many commerically-available and inexpensive IMUs have \textgreater{} 12 hour battery life an onboard storage to record \emph{real-time}, \emph{full-day} data of movements at a sampling rate that is higher than typical video (e.g., 50-100 Hz). Moreover, past work has successfully recorded the rate of leg kicks {[}SMITH CITE{]} and activity intensity {[}CITES{]} in infants and children across an entire day.

The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurement validity is acceptable over long recording periods. IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration. Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space. Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration timeseries) into meaningful categories (e.g., supine, prone, sitting, upright). The difficulty of the classification task depends on the categories of interest. More basic aspects of movement, such as leg kicks or overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity.

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant and be stationary in ``tummy time'' or crawling in myriad ways {[}ADOLPH{]}. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, in a ``W'' position with knees bent. Sitting on a caregivers' lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates still more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life.

Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults {[}CITEs{]}, children {[}CITEs{]}, and infants {[}CITEs{]}. Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest models, that can be later used to predict categories from future IMU data that was not annotated. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion {[}CITE{]} and OTHER EXAMPLES? Our openly-shared dataset {[}CITE{]} contains synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.

REVIEW PAST WORK IN BODY POSITION CLASSIFICATION, END WITH WHAT IS MISSING.
